---
title: "Flap Threshold"
author: "Max Czapanskiy"
date: "`r format(Sys.time(), '%d %B, %Y')`"
knit: (function(inputFile, encoding) { 
         out_dir <- '../analysis/reports';
         out_file <- file.path(dirname(inputFile), out_dir, 'FlapThreshold.html');
         rmarkdown::render(inputFile,
                           encoding = encoding, 
                           output_file = out_file);
       })
output: html_document
---

```{r setup, message = FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	cache = TRUE
)
library(dplyr)
library(lubridate)
library(zoo)
library(ggplot2)
library(foreach)
library(RSQLite)
library(parallel)

# dplyr connections
MHI_db <- src_sqlite('../data/MHI_GPS.sqlite')
metadata_db <- tbl(MHI_db, 'Metadata')
tidytracks_db <- tbl(MHI_db, 'TidyTracks')
acceleration_db <- tbl(MHI_db, 'SegmentACC')

POSIX.origin = ymd('1970-01-01', tz = 'UTC')

load('../data/out/TransitSegments/Segments.RData')
load('../data/out/TransitSegments/ValidSegments.RData')
```

Utility functions
```{r utility}
# This function retrieves segment acceleration data from database
get.acc <- function(segID) {
  # Knitting doesn't play nice with external database connections, so this
  #   is the workaround.
  this.acc_db <- tbl(src_sqlite('../data/MHI_GPS.sqlite'), 'SegmentACC')
  this.acc_db %>%
    filter(SegmentID == segID) %>%
    collect(n = Inf) %>%
    transmute(DeployID,
              SegmentID,
              BurstID,
              ACCTimestampUTC = as.POSIXct(ACCTimestampUTC,
                                           tz = 'UTC',
                                           origin = POSIX.origin),
              X, Y, Z,
              DynX, DynY, DynZ)
}
```

Distribution of Heave Standard Deviation
Sample 1% of the segments, calculate the standard deviation of dynamic heave in 0.5s (10 points) windows. Plot the histogram. Repeat 10 times. Ideally there should be two peaks: a lower one for gliding and a higher one for flapping. If the troughs are in the same place then that's the threshold.
```{r distHeaveSd, warning = FALSE}
# Calculate standard deviation of dynamic heave over 10 point (0.5s) window
sd.win <- 10

# Make ten plots and suppress the output of times
foo <- times(10) %do% {
  # Each plot samples 1% of the segments
  p <- foreach(seg = sample_frac(validSegments, 0.01)$SegmentID,
          .combine = rbind) %do% {
    get.acc(seg) %>%
      # Calculate the sd of DynZ over a sliding window
      mutate(SdDynZ = rollapply(DynZ, sd.win, sd, align = 'center', fill = NA))
  } %>%
    # Create the histogram
    ggplot(aes(x = SdDynZ)) +
    geom_histogram(binwidth = 0.2, boundary = 0) +
    labs(x = 'Standard Deviation of Dynamic Heave (m/s^2)',
         y = 'Count of Acceleration Points',
         title = 'Distribution of Standard Deviation of Dynamic Heave\nIndicates a Threshold at 5m/s^2') +
    xlim(0, 15)
  # ggplot results must be explicitly printed in foreach loops (including times)
  print(p)
}
```

From a visual inspection of the distribution of the standard deviation of dynamic heave I am setting the flap/glide threshold at 5m/s^2.

This chunk has functions for visualizing acceleration by segment.

```{r vizAccFun}
flap.thresh <- 5

# This function plots the heave acceleration bursts from a segment. Color codes 
#   flapping points based on a sliding window of standard deviance and a threshold.
plot.acc <- function(segID) {
  # Retrieve segment acceleration data from database
  # Calculate variance in dynamic heave
  acc.data <- get.acc(segID) %>%
    mutate(SdDynZ = rollapply(DynZ, sd.win, sd, align = 'center', fill = NA),
           FlapGlide = factor(ifelse(SdDynZ >= flap.thresh, 'Flap', 'Glide')))
  
  # Filter segment data
  seg.data <- filter(validSegments, SegmentID == segID)
  
  # Plot acceleration data faceted by burst
  ggplot(acc.data, 
         # Put the acceleration timestamp on the x axis and the dynamic heave on the y
         aes(x = ACCTimestampUTC, y = DynZ)) + 
    # Plot acceleration with both a line and points
    geom_line() + 
    # Points are color coded by flap/glide
    geom_point(aes(color = FlapGlide),
               size = 0.5) + 
    # Facet by burst. Free the x scale so each burst takes up its facet completely.
    facet_wrap(~ BurstID, scales = 'free_x') +
    # Add axis titles, plot title, and a caption with DeployID and segment start time
    labs(x = 'Time (seconds since beginning of burst)', 
         y = 'Dynamic Heave (m/s^2)',
         title = sprintf('Dynamic Heave Along Segment %i', seg.data$SegmentID),
         caption = sprintf('DeployID %i, segment beginning %s',
                           seg.data$DeployID,
                           # %F %R is y-m-d h:m
                           format(seg.data$Start, '%F %R'))) +
    # Burst x-axis labels (time) should be seconds since beginning of the burst. So
    #  we set the x-axis breaks to 1 second and seq_along the breaks to get the labels.
    scale_x_datetime(breaks = function(lim) seq(lim[1], lim[2], by = 1),
                     labels = function(breaks) seq_along(breaks) - 1) +
    # Set the theme to black and white
    theme_bw(base_size = 12) +
    # Remove the legend (use the figure caption for flap/glide coloring)
    theme(legend.position = 'none',
          # Shrink the axis labels
          axis.text = element_text(size = 9),
          # Increase the caption size
          plot.caption = element_text(size = 14),
          # Remove the facet strips
          strip.background = element_blank(),
          strip.text = element_blank())
}
```

Plot the acceleration traces for a subset of the segments. Randomly select 100 segments, plot their acceleration traces, and save to file. Then randomly select 5 more, plot, and print to the report. (foo is used to suppress output of foreach)
```{r plotAcc1}
set.seed(1136)
plot.segs <- sample(validSegments$SegmentID, 100, replace = FALSE)
foo <- foreach(segId = plot.segs) %do% {
  p <- plot.acc(segId)
  plot.path <- '../analysis/figures/SegmentAcceleration/'
  plot.file <- file.path(plot.path, sprintf('%i.png', segId))
  ggsave(plot.file, p, width = 15, height = 8.8, units = 'in')
}
```

```{r plotAcc2}
set.seed(1546)
plot.segs <- sample(validSegments$SegmentID, 5, replace = FALSE)
foo <- foreach(segId = plot.segs) %do% {
  p <- plot.acc(segId)
  print(p)
}
```

Annotate segments with flap ratio
```{r addFlapRatio}
# Functionality for calculating flap ratio per segment based on sd(DynZ) threshold
calculateFlapRatio <- function(segID) {
  acc.data <- get.acc(segID) %>%
    mutate(SdDynZ = rollapply(DynZ, sd.win, sd, align = 'center', fill = NA))
  sum(acc.data$SdDynZ > flap.thresh, na.rm = TRUE) / nrow(acc.data)
}

# Let's go parallel
nCores <- detectCores() - 1
flapCl <- makeCluster(nCores)
clusterExport(flapCl, c('get.acc', 'POSIX.origin', 'sd.win', 'flap.thresh'))
foo <- clusterEvalQ(flapCl, {
  library(dplyr)
  library(zoo)
})
  
# Annotate validSegments with flap ratio
accSegments <- validSegments %>%
  mutate(FlapRatio = parSapply(flapCl, SegmentID, calculateFlapRatio))

stopCluster(flapCl)

# Save results to file
save(accSegments, file = '../data/out/TransitSegments/AccSegments.RData')
```

Flap ratio diagnostic plots
```{r viewFlapRatio}
ggplot(accSegments, aes(x = FlapRatio)) +
  geom_histogram(binwidth = 0.05,
                 boundary = 0) +
  labs(x = 'Flap Ratio',
       y = 'Count of Segments',
       title = 'Distribution of Flap Ratio')

ggplot(accSegments,
       aes(x = Leg,
           y = FlapRatio)) +
  geom_boxplot() +
  labs(x = 'Segment Leg',
       y = 'Flap Ratio',
       title = 'Flap Ratios Are Higher For Outbound Segments')
```
