---
title: "Flap Threshold"
author: "Max Czapanskiy"
date: "`r format(Sys.time(), '%d %B, %Y')`"
knit: (function(inputFile, encoding) { 
         out_dir <- '../analysis/reports';
         out_file <- file.path(dirname(inputFile), out_dir, 'FlapThreshold.html');
         rmarkdown::render(inputFile,
                           encoding = encoding, 
                           output_file = out_file);
       })
output: html_document
---

```{r setup, message = FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	cache = TRUE
)
library(dplyr)
library(lubridate)
library(zoo)
library(ggplot2)
library(foreach)
library(RSQLite)

# dplyr connections
MHI_db <- src_sqlite('../data/MHI_GPS.sqlite')
metadata_db <- tbl(MHI_db, 'Metadata')
tidytracks_db <- tbl(MHI_db, 'TidyTracks')
acceleration_db <- tbl(MHI_db, 'SegmentACC')

POSIX.origin = ymd('1970-01-01', tz = 'UTC')

load('../data/out/TransitSegments/Segments.RData')
load('../data/out/TransitSegments/ValidSegments.RData')
```

Utility functions
```{r utility}
# This function retrieves segment acceleration data from database
get.acc <- function(segID, con = MHI_db) {
  # This allows each call to have its own DB connection, otherwise I get an 
  #   "external pointer is not valid" error in RMarkdown + foreach
  this.acc_db <- tbl(con, 'SegmentACC')
  this.acc_db %>%
    filter(SegmentID == segID) %>%
    collect(n = Inf) %>%
    transmute(DeployID,
              SegmentID,
              BurstID,
              ACCTimestampUTC = as.POSIXct(ACCTimestampUTC,
                                           tz = 'UTC',
                                           origin = POSIX.origin),
              X, Y, Z,
              DynX, DynY, DynZ)
}
```

Distribution of Heave Standard Deviation
Sample 1% of the segments, calculate the standard deviation of dynamic heave in 0.5s (10 points) windows. Plot the histogram. Repeat 10 times. Ideally there should be two peaks: a lower one for gliding and a higher one for flapping. If the troughs are in the same place then that's the threshold.
```{r distHeaveSd, warning = FALSE}
# Calculate standard deviation of dynamic heave over 10 point (0.5s) window
sd.win <- 10

# Make ten plots and suppress the output of times
foo <- times(10) %do% {
  # Each plot samples 1% of the segments
  p <- foreach(seg = sample_frac(validSegments, 0.01)$SegmentID,
          .combine = rbind) %do% {
    get.acc(seg, src_sqlite('../data/MHI_GPS.sqlite')) %>%
      # Calculate the sd of DynZ over a sliding window
      mutate(SdDynZ = rollapply(DynZ, sd.win, sd, align = 'center', fill = NA))
  } %>%
    # Create the histogram
    ggplot(aes(x = SdDynZ)) +
    geom_histogram(binwidth = 0.2, boundary = 0)
  # ggplot results must be explicitly printed in foreach loops (including times)
  print(p)
}
```

From a visual inspection of the distribution of the standard deviation of dynamic heave I am setting the flap/glide threshold at 5m/s^2.

This chunk has functions for visualizing acceleration by segment.

```{r vizAcc}
flap.thresh <- 5

# This function plots the heave acceleration bursts from a segment. Color codes 
#   flapping points based on a sliding window of standard deviance and a threshold.
plot.acc <- function(segID) {
  # Retrieve segment acceleration data from database
  # Calculate variance in dynamic heave
  acc.data <- get.acc(segID) %>%
    mutate(SdDynZ = rollapply(DynZ, window, sd, align = 'center', fill = NA),
           FlapGlide = factor(ifelse(SdDynZ >= flap.thresh, 'Flap', 'Glide')))
  
  # Filter segment data
  seg.data <- filter(validSegments, SegmentID == segID)
  
  # Plot acceleration data faceted by burst
  ggplot(acc.data, aes(x = ACCTimestampUTC, y = DynZ)) + 
    geom_line() + 
    geom_point(size = 0.5) + 
    facet_wrap(~ BurstID, scales = 'free_x') +
    labs(x = 'Time', 
         y = 'Dynamic Heave (m/s^2)',
         title = sprintf('Dynamic Heave Along Segment %i', seg.data$SegmentID),
         caption = sprintf('DeployID %i, segment beginning %s',
                           seg.data$DeployID,
                           format(seg.data$Start, '%F %R')))
}
```
