---
title: "Flap Threshold"
author: "Max Czapanskiy"
date: "`r format(Sys.time(), '%d %B, %Y')`"
knit: (function(inputFile, encoding) { 
         out_dir <- '../analysis/reports';
         out_file <- file.path(dirname(inputFile), out_dir, 'FlapThreshold.html');
         rmarkdown::render(inputFile,
                           encoding = encoding, 
                           output_file = out_file);
       })
output: html_document
---

```{r setup, message = FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	cache = TRUE
)
library(dplyr)
library(lubridate)
library(zoo)
library(ggplot2)
library(foreach)
library(RSQLite)
library(parallel)

# dplyr connections
MHI_db <- src_sqlite('../data/MHI_GPS.sqlite')
metadata_db <- tbl(MHI_db, 'Metadata')
tidytracks_db <- tbl(MHI_db, 'TidyTracks')
acceleration_db <- tbl(MHI_db, 'SegmentACC')

POSIX.origin = ymd('1970-01-01', tz = 'UTC')

load('../data/out/TransitSegments/Segments.RData')
load('../data/out/TransitSegments/ValidSegments.RData')
```

Utility functions
```{r utility}
# This function retrieves segment acceleration data from database
get.acc <- function(segID) {
  # Knitting doesn't play nice with external database connections, so this
  #   is the workaround.
  mhi_db <- DBI::dbConnect(RSQLite::SQLite(), dbname = '../data/MHI_GPS.sqlite')
  this.acc_db <- tbl(mhi_db, 'SegmentACC')
  result <- this.acc_db %>%
    filter(SegmentID == segID) %>%
    collect(n = Inf) %>%
    transmute(DeployID,
              SegmentID,
              BurstID,
              ACCTimestampUTC = as.POSIXct(ACCTimestampUTC,
                                           tz = 'UTC',
                                           origin = POSIX.origin),
              X, Y, Z,
              DynX, DynY, DynZ,
              ODBA)
  DBI::dbDisconnect(mhi_db)
  result
}
```

Distribution of Heave Standard Deviation
Sample 1% of the segments, calculate the standard deviation of dynamic heave in 0.5s (10 points) windows. Plot the histogram. Repeat 10 times. Ideally there should be two peaks: a lower one for gliding and a higher one for flapping. If the troughs are in the same place then that's the threshold.
```{r distHeaveSd, warning = FALSE, eval = FALSE}
# Calculate standard deviation of dynamic heave over various windows
foreach(sd.win = 5:15) %do% {
  validSegments %>%
    group_by(DeployID) %>%
    do({
      n_seg <- min(nrow(.), 100)
      sdDynZdist <- sample_n(., n_seg, replace = FALSE) %>%
        rowwise %>%
        do({
          get.acc(.$SegmentID) %>%
            mutate(SdDynZ = rollapply(DynZ, sd.win, sd, align = 'center', fill = NA))
        })
      p <- ggplot(sdDynZdist, aes(x = SdDynZ)) +
        geom_histogram(binwidth = 0.2, boundary = 0) +
        labs(x = 'Standard Deviation of Dynamic Heave (m/s^2)',
             y = 'Count of Acceleration Points',
             title = 'Distribution of Standard Deviation of Dynamic Heave',
             caption = sprintf('Window of %d points\nDeployID %s', sd.win, first(.$DeployID))) +
        xlim(0, 15)
      ggsave(sprintf('../analysis/figures/SegmentAcceleration/SdWin/%s_%i.png', 
                     first(.$DeployID), 
                     sd.win),
             p, width = 15, height = 8.8, units = 'in')
      data.frame(DeployID = first(.$DeployID, sdWin = sd.win))
    })
}

# Optimal window appears to be 6 points (0.3 seconds)
sd.win <- 6
```

A visual inspection of the distribution of the standard deviation of dynamic shows that a window of six points has the clearest bimodality and there's a flap/glide threshold at 5.5m/s^2.

This chunk has functions for visualizing acceleration by segment.

```{r vizAccFun}
flap.thresh <- 5.5

# This function plots the heave acceleration bursts from a segment. Color codes 
#   flapping points based on a sliding window of standard deviance and a threshold.
plot.acc <- function(segID) {
  # Retrieve segment acceleration data from database
  # Calculate variance in dynamic heave
  acc.data <- get.acc(segID) %>%
    mutate(SdDynZ = rollapply(DynZ, sd.win, sd, align = 'center', fill = NA),
           FlapGlide = factor(ifelse(SdDynZ >= flap.thresh, 'Flap', 'Glide')))
  
  # Filter segment data
  seg.data <- filter(validSegments, SegmentID == segID)
  
  # Plot acceleration data faceted by burst
  ggplot(acc.data, 
         # Put the acceleration timestamp on the x axis and the dynamic heave on the y
         aes(x = ACCTimestampUTC, y = DynZ)) + 
    # Plot acceleration with both a line and points
    geom_line() + 
    # Points are color coded by flap/glide
    geom_point(aes(color = FlapGlide),
               size = 0.5) + 
    # Facet by burst. Free the x scale so each burst takes up its facet completely.
    facet_wrap(~ BurstID, scales = 'free_x') +
    # Add axis titles, plot title, and a caption with DeployID and segment start time
    labs(x = 'Time (seconds since beginning of burst)', 
         y = 'Dynamic Heave (m/s^2)',
         title = sprintf('Dynamic Heave Along Segment %i', seg.data$SegmentID),
         caption = sprintf('DeployID %i, segment beginning %s',
                           seg.data$DeployID,
                           # %F %R is y-m-d h:m
                           format(seg.data$Start, '%F %R'))) +
    # Burst x-axis labels (time) should be seconds since beginning of the burst. So
    #  we set the x-axis breaks to 1 second and seq_along the breaks to get the labels.
    scale_x_datetime(breaks = function(lim) seq(lim[1], lim[2], by = 1),
                     labels = function(breaks) seq_along(breaks) - 1) +
    # Set the theme to black and white
    theme_bw(base_size = 12) +
    # Remove the legend (use the figure caption for flap/glide coloring)
    theme(legend.position = 'none',
          # Shrink the axis labels
          axis.text = element_text(size = 9),
          # Increase the caption size
          plot.caption = element_text(size = 14),
          # Remove the facet strips
          strip.background = element_blank(),
          strip.text = element_blank())
}
```

Plot the acceleration traces for a subset of the segments. Randomly select 100 segments, plot their acceleration traces, and save to file. Then randomly select 5 more, plot, and print to the report. (foo is used to suppress output of foreach)
```{r plotAcc1}
set.seed(1136)
plot.segs <- sample(validSegments$SegmentID, 100, replace = FALSE)
foo <- foreach(segId = plot.segs) %do% {
  p <- plot.acc(segId)
  plot.path <- '../analysis/figures/SegmentAcceleration/'
  plot.file <- file.path(plot.path, sprintf('%i.png', segId))
  ggsave(plot.file, p, width = 15, height = 8.8, units = 'in')
}
```

```{r plotAcc2}
set.seed(1546)
plot.segs <- sample(validSegments$SegmentID, 5, replace = FALSE)
foo <- foreach(segId = plot.segs) %do% {
  p <- plot.acc(segId)
  print(p)
}
```

Annotate segments with flap ratio
Note: This seems to make RMarkdown hang. I'm leaving the code in here but it doesn't run. I run it separately and save the results to AccSegments.RData.
```{r addFlapRatio, eval = FALSE}
# Functionality for calculating flap ratio per segment based on sd(DynZ) threshold
calculateFlapRatio <- function(segID) {
  acc.data <- get.acc(segID) %>%
    mutate(SdDynZ = rollapply(DynZ, sd.win, sd, align = 'center', fill = NA))
  sum(acc.data$SdDynZ > flap.thresh, na.rm = TRUE) / nrow(acc.data)
}
calculateODBA <- function(segID) {
  acc.data <- get.acc(segID)
  mean(acc.data$ODBA, na.rm = TRUE)
}

# Let's go parallel
nCores <- detectCores() - 1
flapCl <- makeCluster(nCores)
clusterExport(flapCl, c('get.acc', 'POSIX.origin', 'sd.win', 'flap.thresh'))
foo <- clusterEvalQ(flapCl, {
  library(dplyr)
  library(zoo)
})
  
# Annotate validSegments with flap ratio
accSegments <- validSegments %>%
  mutate(FlapRatio = parSapply(flapCl, SegmentID, calculateFlapRatio),
         MeanODBA = parSapply(flapCl, SegmentID, calculateODBA))

stopCluster(flapCl)

# Save results to file
save(accSegments, file = '../data/out/TransitSegments/AccSegments.RData')
```

Flap ratio diagnostic plots
```{r viewFlapRatio}
load('../data/out/TransitSegments/AccSegments.RData')

ggplot(accSegments, aes(x = FlapRatio)) +
  geom_histogram(binwidth = 0.05,
                 boundary = 0) +
  labs(x = 'Flap Ratio',
       y = 'Count of Segments',
       title = 'Distribution of Flap Ratio')

ggplot(accSegments,
       aes(x = Leg,
           y = FlapRatio)) +
  geom_boxplot() +
  labs(x = 'Segment Leg',
       y = 'Flap Ratio',
       title = 'Flap Ratios Are Higher For Outbound Segments')
```
