---
title: "Flight Model"
author: "Max Czapanskiy"
date: "`r format(Sys.time(), '%d %B, %Y')`"
knit: (function(inputFile, encoding) { 
         out_dir <- '../analysis/reports';
         out_file <- file.path(dirname(inputFile), out_dir, 'FlightModelGAMM2.html');
         rmarkdown::render(inputFile,
                           encoding = encoding, 
                           output_file = out_file);
       })
output: html_document
---
  
```{r setup, message = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  cache = TRUE
)
library(tidyverse)
library(lubridate)
library(RSQLite)
library(knitr)
library(akima)
library(rgdal)
library(geosphere)
library(spdep)
library(mosaic)
library(rlang)
library(mgcv)
library(lme4)
library(grid)
library(gridExtra)
library(ggthemes)
library(visreg)

# dplyr connections
MHI_db <- src_sqlite('../data/MHI_GPS.sqlite')
metadata_db <- tbl(MHI_db, 'Metadata')
tidytracks_db <- tbl(MHI_db, 'TidyTracks')
acceleration_db <- tbl(MHI_db, 'SegmentACC')

POSIX.origin = ymd('1970-01-01', tz = 'UTC')

# Load data from earlier reports
load('../data/out/TransitSegments/Segments.RData')
load('../data/out/TransitSegments/ValidSegments.RData')
load('../data/out/TransitSegments/AccSegments.RData')
load('../data/out/TransitSegments/WindSegments.RData')

select <- dplyr::select
```

How do seabirds modulate flight behavior in response to wind? This report seeks to answer that question by modeling the flap ratio as a function of tail- (`T`) and crosswinds (`C`) with individual bird (`B`) and trip (`R`) as random effects.

First up: segment sampling. Randomly select both an outbound and inbound segment from each eligible trip (i.e. trips _both_ outbound and inbound segments).
```{r sample_segments}
# Which trips have both outbound and inbound legs?
bothLegs <- windSegments %>%
  group_by(DeployID, TripID, Leg) %>%
  summarize(N = n()) %>%
  ungroup %>%
  spread(Leg, N) %>%
  mutate(Both = !is.na(Inbound) & !is.na(Outbound)) %>%
  filter(Both)

set.seed(211)
segSample <- windSegments %>%
  semi_join(bothLegs, by = 'TripID') %>%
  group_by(TripID, Leg) %>%
  sample_n(1) %>%
  ungroup %>%
  mutate(AirSpd = sqrt((meanSpd - Tailwind)^2 + Crosswind^2),
         Leg = factor(Leg)) 

save(segSample, file = '../data/out/TransitSegments/SegSample.RData')
```

Let's look at the distribution of segments across deployments, time, and location.
```{r view_segments}
# Across deployments
ggplot(segSample, aes(x = DeployID)) +
  geom_bar() +
  labs(title = 'Segment sample: 14-48 segments from 12 individuals')

# Through time
ggplot(segSample, aes(x = Start, y = DeployID)) +
  geom_point() +
  labs(x = 'Date of Segment',
       title = 'Distribution of segments between May 28 and July 17')

# In space
ColLon <- -159.3997
ColLat <- 22.22888
LonExtent <- c(min(ColLon, segSample$Longitude), max(ColLon, segSample$Longitude))
LatExtent <- c(min(ColLat, segSample$Latitude), max(ColLat, segSample$Latitude))
hi_shp <- readOGR('../data/coastline/ne_10m_coastline/', 'ne_10m_coastline')
hi_df <- fortify(hi_shp) %>%
  filter(between(long, LonExtent[1], LonExtent[2]),
         between(lat, LatExtent[1], LatExtent[2]))
ggplot(segSample) +
  geom_path(data = hi_df, aes(x = long, y = lat, group = group)) +
  geom_point(aes(x = Longitude, y = Latitude, color = Start)) +
  coord_map() +
  theme(legend.position = 'bottom')
```

```{r plot_distributions}
angle_breaks <- pi/4 * 0:4
angle_lbls <- c('0', 
                expr(pi/4),
                expr(pi/2),
                expr(3*pi/4),
                expr(pi))
angle_scale <- scale_x_continuous(breaks = angle_breaks,
                                  labels = angle_lbls,
                                  name = '(Tailwind)  <---  Wind Angle  --->  (Headwind)')
p3 <- ggplot(segSample, aes(x = WindAngle)) +
  geom_histogram(binwidth = pi/16, boundary = 0) + 
  angle_scale +
  labs(y = 'frequency') +
  theme_few(base_size = 9)
png('../analysis/figures/Fig3.png', width = 85, height = 85, units = 'mm', res = 300)
print(p3)
dev.off()
  
ggplot(segSample, aes(x = WindSpd)) +
  geom_histogram(binwidth = 0.5, boundary = 0) +
  labs(x = 'Wind Speed (m/s)')

mean(segSample$WindSpd)
sd(segSample$WindSpd)

p4a <- ggplot(segSample, aes(x = FlapRatio)) +
  geom_histogram(binwidth = 0.02, boundary = 0) +
  annotate('text', x = 0.05, y = 30, label = 'bold(a)', size = 8, parse = TRUE) +
  scale_x_continuous(breaks = 1:6 / 10,
                     labels = paste0(10 * 1:6, '%')) +
  labs(x = 'Flap Ratio',
       y = 'frequency') +
  theme_few(base_size = 9)

p4b <- ggplot(segSample, aes(x = MeanODBA)) +
  geom_histogram(binwidth = 0.2, boundary = 0) +
  annotate('text', x = 4.5, y = 28, label = 'bold(b)', size = 8, parse = TRUE) +
  labs(x = 'ODBA (m/s^2)',
       y = '') +
  theme_few(base_size = 9)

png('../analysis/figures/Fig4.png', width = 170, height = 85, units = 'mm', res = 300)
pushViewport(viewport(layout = grid.layout(1, 2)))
print(p4a, vp = viewport(layout.pos.row = 1, layout.pos.col = 1))
print(p4b, vp = viewport(layout.pos.row = 1, layout.pos.col = 2))
dev.off()

quantile(segSample$FlapRatio)
mean(segSample$FlapRatio)
sd(segSample$FlapRatio)
quantile(segSample$MeanODBA)
mean(segSample$MeanODBA)
sd(segSample$MeanODBA)

ggplot(segSample, aes(x = meanSpd)) +
  geom_histogram(binwidth = 1, boundary = 0) +
  labs(x = 'Ground speed (m/s)')

quantile(segSample$meanSpd)
mean(segSample$meanSpd)
sd(segSample$meanSpd)
```

Now a look at the response varibles (F, O) in response to the predictor variables (T, C, A, M)
```{r plot_vars_FV}
# F ~ T
ggplot(segSample, aes(x = Tailwind, y = FlapRatio)) +
  geom_point() +
  geom_smooth(method = 'loess') +
  labs(x = 'Tailwind (m/s)',
       y = 'Flap Ratio') +
  theme(legend.position = 'bottom')

# F ~ C
ggplot(segSample, aes(x = Crosswind, y = FlapRatio, color = factor(Tailwind > 0, labels = c('Tailwind', 'Headwind')))) +
  geom_point() +
  geom_smooth(method = 'loess') +
  labs(x = 'Crosswind (m/s)',
       y = 'Flap Ratio',
       color = 'Tailwind/Headwind') +
  theme(legend.position = 'bottom')

# F ~ T:C
ggplot(segSample, aes(x = Tailwind * Crosswind, y = FlapRatio)) +
  geom_point() +
  geom_smooth(method = 'loess') +
  labs(x = 'Tailwind * Crosswind (m^2/s^2)',
       y = 'Flap Ratio',
       title = 'Flap ratio relationship to tailwind:crosswind') +
  theme(legend.position = 'bottom')

Cross <- seq(0, 11, by = 1); Cstep <- Cross[2] - Cross[1];
Tail <- seq(-11, 11, by = 1); Tstep <- Tail[2] - Tail[1];
expand.grid(Crosswind = Cross, Tailwind = Tail) %>%
  mutate(FlapRatio = mapply(Tailwind, Crosswind,
                            FUN = function(T1, C1) {
                              points <- filter(segSample,
                                               Tailwind >= T1,
                                               Tailwind < T1 + Tstep,
                                               Crosswind >= C1,
                                               Crosswind < C1 + Cstep)
                              if(nrow(points) == 0) 
                                NA 
                              else
                                mean(points$FlapRatio, na.rm = TRUE)
                            })) %>% 
  ggplot(aes(x = Tailwind, y = Crosswind, fill = FlapRatio)) +
  geom_raster() +
  scale_fill_distiller(palette = 'RdYlBu')
```

```{r plot_vars_FS}
# F ~ A
p5 <- ggplot(segSample, aes(x = WindAngle, y = FlapRatio)) +
  geom_point(color = 'gray') +
  geom_smooth(method = 'loess', se = FALSE, color = 'black') +
  labs(x = 'Wind Angle',
       y = 'Flap Ratio') +
  scale_y_continuous(breaks = 1:5 / 10,
                     labels = paste0(10 * 1:5, '%')) +
  angle_scale +
  theme_few(base_size = 9)

print(p5)

png('../analysis/figures/Fig5.png', width = 85, height = 85, units = 'mm', res = 300)
print(p5)
dev.off()

# F ~ M
ggplot(segSample, aes(x = WindSpd, y = FlapRatio, color = WindAngle)) +
  geom_point() +
  geom_smooth(method = 'loess') +
  labs(x = 'Wind Magnitude (m/s)',
       y = 'FlapRatio')

# F ~ A * M
A <- seq(0, pi, length.out = 10)
Astep <- A[2] - A[1]
M <- seq(0, 12, by = 1)
Mstep <- M[2] - M[1];
expand.grid(WindAngle = A, WindSpd = M) %>%
  mutate(FlapRatio = mapply(WindAngle, WindSpd,
                            FUN = function(A1, M1) {
                              points <- filter(segSample,
                                               WindAngle >= A1,
                                               WindAngle < A1 + Astep,
                                               WindSpd >= M1,
                                               WindSpd < M1 + Mstep)
                              if(nrow(points) == 0) 
                                NA 
                              else
                                mean(points$FlapRatio, na.rm = TRUE)
                            })) %>% 
  ggplot(aes(x = WindAngle, y = WindSpd, fill = FlapRatio)) +
  geom_raster() +
  scale_fill_distiller(palette = 'RdYlBu') + 
  scale_x_continuous(breaks = pi/4 *0:4,
                     labels = c('0', 'pi/4', 'pi/2', '3pi/4', 'pi'),
                     name = 'Tailwind  <---  Wind Angle  --->  Headwind')
```

```{r plot_vars_OV}
# O ~ T
ggplot(segSample, aes(x = Tailwind, y = MeanODBA)) +
  geom_point() +
  geom_smooth(method = 'loess') +
  labs(x = 'Tailwind (m/s)',
       y = 'ODBA (m/s^2)') +
  theme(legend.position = 'bottom')

# O ~ C
ggplot(segSample, aes(x = Crosswind, y = MeanODBA, color = factor(Tailwind > 0, labels = c('Tailwind', 'Headwind')))) +
  geom_point() +
  geom_smooth(method = 'loess') +
  labs(x = 'Crosswind (m/s)',
       y = 'ODBA (m/s^2)',
       color = 'Tailwind/Headwind') +
  theme(legend.position = 'bottom')

# O ~ T:C
ggplot(segSample, aes(x = Tailwind * Crosswind, y = MeanODBA)) +
  geom_point() +
  geom_smooth(method = 'loess') +
  labs(x = 'Tailwind * Crosswind (m^2/s^2)',
       y = 'ODBA (m/s^2)') +
  theme(legend.position = 'bottom')

Cross <- seq(0, 11, by = 1); Cstep <- Cross[2] - Cross[1];
Tail <- seq(-11, 11, by = 1); Tstep <- Tail[2] - Tail[1];
expand.grid(Crosswind = Cross, Tailwind = Tail) %>%
  mutate(MeanODBA = mapply(Tailwind, Crosswind,
                            FUN = function(T1, C1) {
                              points <- filter(segSample,
                                               Tailwind >= T1,
                                               Tailwind < T1 + Tstep,
                                               Crosswind >= C1,
                                               Crosswind < C1 + Cstep)
                              if(nrow(points) == 0) 
                                NA 
                              else
                                mean(points$MeanODBA, na.rm = TRUE)
                            })) %>% 
  ggplot(aes(x = Tailwind, y = Crosswind, fill = MeanODBA)) +
  geom_raster() +
  scale_fill_distiller(palette = 'RdYlBu')
```

```{r plot_vars_OS}
# O ~ A
ggplot(segSample, aes(x = WindAngle, y = MeanODBA)) +
  geom_point() +
  geom_smooth(method = 'loess') +
  labs(x = 'Wind Angle',
       y = 'ODBA (m/s^2)') +
  scale_x_continuous(breaks = pi/4 *0:4,
                     labels = c('0', 'pi/4', 'pi/2', '3pi/4', 'pi'),
                     name = 'Tailwind  <---  Wind Angle  --->  Headwind')

# O ~ M
ggplot(segSample, aes(x = WindSpd, y = MeanODBA, color = WindAngle)) +
  geom_point() +
  geom_smooth(method = 'loess') +
  labs(x = 'Wind Magnitude (m/s)',
       y = 'ODBA (m/s^2)')

# O ~ A * M
A <- seq(0, pi, length.out = 10)
Astep <- A[2] - A[1]
M <- seq(0, 12, by = 1)
Mstep <- M[2] - M[1];
expand.grid(WindAngle = A, WindSpd = M) %>%
  mutate(ODBA = mapply(WindAngle, WindSpd,
                            FUN = function(A1, M1) {
                              points <- filter(segSample,
                                               WindAngle >= A1,
                                               WindAngle < A1 + Astep,
                                               WindSpd >= M1,
                                               WindSpd < M1 + Mstep)
                              if(nrow(points) == 0) 
                                NA 
                              else
                                mean(points$MeanODBA, na.rm = TRUE)
                            })) %>% 
  ggplot(aes(x = WindAngle, y = WindSpd, fill = ODBA)) +
  geom_raster() +
  scale_fill_distiller(palette = 'RdYlBu') + 
  scale_x_continuous(breaks = pi/4 *0:4,
                     labels = c('0', 'pi/4', 'pi/2', '3pi/4', 'pi'),
                     name = 'Tailwind  <---  Wind Angle  --->  Headwind')
```


# Model Fitting
```{r flight_gamms}
ftc <- gam(FlapRatio ~ s(Tailwind) + s(Crosswind) + ti(Tailwind, Crosswind) + s(DeployID, bs = 're'), family = betar(), data = segSample)
fam <- gam(FlapRatio ~ s(WindAngle) + s(WindSpd) + ti(WindAngle, WindSpd) + s(DeployID, bs = 're'), family = betar(), data = segSample)
otc <- gam(MeanODBA ~ s(Tailwind) + s(Crosswind) + ti(Tailwind, Crosswind) + s(DeployID, bs = 're'), family = gaussian(), data = segSample)
oam <- gam(MeanODBA ~ s(WindAngle) + s(WindSpd) + ti(WindAngle, WindSpd) + s(DeployID, bs = 're'), family = gaussian(), data = segSample)

summary(ftc)
summary(fam)
summary(otc)
summary(oam)

# Visualize both on a angle/speed plot
grain <- 15
model_predictions <- expand.grid(WindAngle = seq(0, pi, length.out = grain),
                                 WindSpd = seq(4, 12, length.out = grain),
                                 DeployID = 1145) %>%
  mutate(Tailwind = WindSpd * cos(WindAngle),
         Crosswind = WindSpd * sin(WindAngle)) %>%
  mutate(FTC = predict(ftc, newdata = ., 
                       exclude = 's(DeployID)', type = 'response'),
         FAM = predict(fam, newdata = ., 
                       exclude = 's(DeployID)', type = 'response'),
         OTC = predict(otc, newdata = ., 
                       exclude = 's(DeployID)', type = 'response'),
         OAM = predict(oam, newdata = ., 
                       exclude = 's(DeployID)', type = 'response'))
AS_kd <- ks::kde(select(segSample, WindAngle, WindSpd), compute.cont = TRUE)
fence <- function(x, lower, upper) {
    ifelse(x < lower, lower, ifelse(x > upper, upper, x))
}
cont_95 <- with(AS_kd, contourLines(x = eval.points[[1]], 
                                    y = eval.points[[2]],
                                    z = estimate, 
                                    levels = cont["5%"])[[1]]) %>%
  data.frame %>%
  mutate(x = fence(x, 
                   min(model_predictions$WindAngle), 
                   max(model_predictions$WindAngle)),
         y = fence(y, 
                   min(model_predictions$WindSpd), 
                   max(model_predictions$WindSpd)))

# Color blind safe palette
cbs_blue <- '#56B4E9'
cbs_orange <- '#E69F00'
cbs_green <- '#009E73'

p7a <- ggplot(model_predictions, aes(WindAngle, WindSpd)) +
  geom_raster(aes(fill = FTC)) +
  stat_contour(aes(z = FTC), color = cbs_green, binwidth = 0.02) + 
  geom_path(data = cont_95, aes(x, y), linetype = 'dashed', alpha = 0.5) +
  scale_fill_gradient2(low = cbs_blue, 
                       high = cbs_orange, 
                       midpoint = (0.18 + 0.39) / 2,
                       limits = c(0.18, 0.39),
                       labels = function(x) sprintf('%i%%', x * 100)) +
  angle_scale +
  theme_few(base_size = 9) +
  theme(legend.position = 'none', axis.title.y = element_blank()) +
  labs(y = 'Wind Speed (m/s)',
       fill = 'Flap\nRatio')

p7b <- ggplot(model_predictions, aes(WindAngle, WindSpd)) +
  geom_raster(aes(fill = FAM)) +
  stat_contour(aes(z = FAM), color = cbs_green, binwidth = 0.02) + 
  geom_path(data = cont_95, aes(x, y), linetype = 'dashed', alpha = 0.5) +
  scale_fill_gradient2(low = cbs_blue, 
                       high = cbs_orange, 
                       midpoint = (0.18 + 0.39) / 2,
                       limits = c(0.18, 0.39),
                       labels = function(x) sprintf('%i%%', x * 100)) +
  angle_scale +
  theme_few(base_size = 9) +
  theme(legend.position = 'none', axis.title.y = element_blank())

# Axis title and fill guide for 7a & b
p7ab <- ggplot(model_predictions, aes(WindAngle, WindSpd)) +
  geom_raster(aes(fill = FAM)) +
  stat_contour(aes(z = FAM), color = cbs_green, binwidth = 0.02) + 
  geom_path(data = cont_95, aes(x, y), linetype = 'dashed', alpha = 0.5) +
  scale_fill_gradient2(low = cbs_blue, 
                       high = cbs_orange, 
                       midpoint = (0.18 + 0.39) / 2,
                       limits = c(0.18, 0.39),
                       labels = function(x) sprintf('%i%%', x * 100)) +
  angle_scale +
  theme_few(base_size = 9) +
  theme(legend.position = 'right') +
  labs(y = 'Wind Speed (m/s)',
       fill = 'Flap\nRatio')

png('../analysis/figures/Fig7a.png', width = 75, height = 65, units = 'mm', res = 300)
print(p7a)
dev.off()
png('../analysis/figures/Fig7b.png', width = 75, height = 65, units = 'mm', res = 300)
print(p7b)
dev.off()
png('../analysis/figures/Fig7ab.png', width = 75, height = 65, units = 'mm', res = 300)
print(p7ab)
dev.off()

p7c <- ggplot(model_predictions, aes(WindAngle, WindSpd)) +
  geom_raster(aes(fill = OTC)) +
  stat_contour(aes(z = OTC), color = cbs_green, binwidth = 0.5) + 
  geom_path(data = cont_95, aes(x, y), linetype = 'dashed', alpha = 0.5) +
  scale_fill_gradient2(low = cbs_blue, 
                       high = cbs_orange, 
                       midpoint = (6 + 9.2) / 2,
                       limits = c(6, 9.2)) +
  angle_scale +
  theme_few(base_size = 9) +
  theme(legend.position = 'none', axis.title.y = element_blank()) +
  labs(y = 'Wind Speed (m/s)',
       fill = 'Flap\nRatio')

p7d <- ggplot(model_predictions, aes(WindAngle, WindSpd)) +
  geom_raster(aes(fill = OAM)) +
  stat_contour(aes(z = OAM), color = cbs_green, binwidth = 0.5) + 
  geom_path(data = cont_95, aes(x, y), linetype = 'dashed', alpha = 0.5) +
  scale_fill_gradient2(low = cbs_blue, 
                       high = cbs_orange, 
                       midpoint = (6 + 9.2) / 2,
                       limits = c(6, 9.2)) +
  angle_scale +
  theme_few(base_size = 9) +
  theme(legend.position = 'none', axis.title.y = element_blank())

# Axis title and fill guide for 7c & d
p7cd <- ggplot(model_predictions, aes(WindAngle, WindSpd)) +
  geom_raster(aes(fill = OAM)) +
  stat_contour(aes(z = OAM), color = cbs_green, binwidth = 0.5) + 
  geom_path(data = cont_95, aes(x, y), linetype = 'dashed', alpha = 0.5) +
  scale_fill_gradient2(low = cbs_blue, 
                       high = cbs_orange, 
                       midpoint = (6 + 9.2) / 2,
                       limits = c(6, 9.2)) +
  angle_scale +
  theme_few(base_size = 9) +
  theme(legend.position = 'right') +
  labs(y = 'Wind Speed (m/s)',
       fill = 'ODBA\n(m/s^2)')

png('../analysis/figures/Fig7c.png', width = 75, height = 65, units = 'mm', res = 300)
print(p7c)
dev.off()
png('../analysis/figures/Fig7d.png', width = 75, height = 65, units = 'mm', res = 300)
print(p7d)
dev.off()
png('../analysis/figures/Fig7cd.png', width = 75, height = 65, units = 'mm', res = 300)
print(p7cd)
dev.off()
```

# Model Validation
Use cross validation to see which model (FTC, FAM) is best

```{r model_validation_F}
cross_validate <- function() {
  frac <- 0.8

  fit_data <- sample_frac(segSample, size = frac)
  check_data <- anti_join(segSample, fit_data, by = 'SegmentID')
  
  tryCatch({
    ftc_frac <- gam(FlapRatio ~ s(Tailwind) + s(Crosswind) + ti(Tailwind, Crosswind) + s(DeployID, bs = 're'), family = betar(), data = fit_data)
  fam_frac <- gam(FlapRatio ~ s(WindAngle) + s(WindSpd) + ti(WindAngle, WindSpd) + s(DeployID, bs = 're'), family = betar(), data = fit_data)
  
  pred_data <- check_data %>%
    mutate(ftc_pred = predict(ftc_frac, ., type = 'response'),
           fam_pred = predict(fam_frac, ., type = 'response'),
           ftc_resid = ftc_pred - FlapRatio,
           fam_resid = fam_pred - FlapRatio)
  
  pred_data %>%
    summarize(ftc_rmse = sqrt(1/n() * sum(ftc_resid^2)),
              fam_rmse = sqrt(1/n() * sum(fam_resid^2)))
  }, error = function(e) data.frame(otc_rmse = NA, oam_rmse = NA))
}

cv_results <- replicate(1e4, cross_validate(), simplify = TRUE) %>%
  t %>%
  data.frame %>%
  mutate_all(as.numeric)

cv_results %>% 
  gather(model, rmse, ftc_rmse, fam_rmse) %>% 
  ggplot(aes(x = rmse, color = model)) +
  geom_density()

t.test(cv_results$ftc_rmse, cv_results$fam_rmse)
```
Use cross validation to see which model (OTC, OAM) is best

```{r model_validation_O}
cross_validate <- function() {
  frac <- 0.8

  fit_data <- sample_frac(segSample, size = frac)
  check_data <- anti_join(segSample, fit_data, by = 'SegmentID')
  
  tryCatch({
    otc_frac <- gam(MeanODBA ~ s(Tailwind) + s(Crosswind) + ti(Tailwind, Crosswind) + s(DeployID, bs = 're'), family = gaussian(), data = fit_data)
    oam_frac <- gam(MeanODBA ~ s(WindAngle) + s(WindSpd) + ti(WindAngle, WindSpd) + s(DeployID, bs = 're'), family = gaussian(), data = fit_data) 
    
    pred_data <- check_data %>%
      mutate(otc_pred = predict(otc_frac, ., type = 'response'),
             oam_pred = predict(oam_frac, ., type = 'response'),
             otc_resid = otc_pred - MeanODBA,
             oam_resid = oam_pred - MeanODBA)
    
    pred_data %>%
      summarize(otc_rmse = sqrt(1/n() * sum(otc_resid^2)),
                oam_rmse = sqrt(1/n() * sum(oam_resid^2)))
  }, error = function(e) data.frame(otc_rmse = NA, oam_rmse = NA))
}

cv_results <- replicate(1e4, cross_validate(), simplify = TRUE) %>%
  t %>%
  data.frame %>%
  mutate_all(as.numeric)

cv_results %>% 
  gather(model, rmse, otc_rmse, oam_rmse) %>% 
  ggplot(aes(x = rmse, color = model)) +
  geom_density()

t.test(cv_results$otc_rmse, cv_results$oam_rmse)
```

```{r diagnostics}
partial_plot <- function(model, predictor, x_scale, y_scale) {
  visreg(model, 
         predictor, 
         type = 'conditional', 
         gg = TRUE, 
         scale = 'response', 
         rug = 1,
         line = list(col = 'black')) + 
    x_scale +
    y_scale +
    theme_few(base_size = 8)
}

T_scale <- scale_x_continuous('Tailwind (m/s)')
C_scale <- scale_x_continuous('Crosswind (m/s)')
A_scale <- scale_x_continuous('Wind Angle',
                              breaks = angle_breaks,
                              labels = angle_lbls)
M_scale <- scale_x_continuous('Wind Speed (m/s)')
F_scale <- scale_y_continuous('Flap Ratio',
                     limits = c(0.2, 0.45),
                     labels = function(x) sprintf('%i%%', x * 100))
O_scale <- scale_y_continuous('ODBA (m/s^2)',
                     limits = c(5, 9))

partials <- mapply(partial_plot,
                   list(ftc, ftc, fam, fam, otc, otc, oam, oam),
                   list('Tailwind', 'Crosswind', 'WindAngle', 'WindSpd',
                        'Tailwind', 'Crosswind', 'WindAngle', 'WindSpd'),
                   list(T_scale, C_scale, A_scale, M_scale,
                        T_scale, C_scale, A_scale, M_scale),
                   list(F_scale, F_scale, F_scale, F_scale, 
                        O_scale, O_scale, O_scale, O_scale),
                   SIMPLIFY = FALSE)

p8 <- arrangeGrob(grobs = partials, nrow = 2)
grid.arrange(p8)
png('../analysis/figures/Fig8.png', width = 170, height = 110, units = 'mm', res = 300)
grid.arrange(p8)
dev.off()
```

```{r speed_models}
p6 <- ggplot(segSample, aes(Tailwind, meanSpd)) +
  geom_point(color = 'gray') +
  geom_smooth(method = 'loess', se = FALSE, color = 'black') +
  labs(x = 'Tailwind Speed (m/s)',
       y = 'Ground Speed (m/s)') +
  theme_few(base_size = 9)

print(p6)

png('../analysis/figures/Fig6.png', width = 85, height = 85, units = 'mm', res = 300)
print(p6)
dev.off()

SgTCcand <- list(lmer(meanSpd ~ poly(Tailwind, 2) * Crosswind + (1|DeployID),
                      data = segSample),
                 lmer(meanSpd ~ poly(Tailwind, 2) * Crosswind + (poly(Tailwind, 2)|DeployID),
                      data = segSample),
                 lmer(meanSpd ~ poly(Tailwind, 2) + Crosswind + (1|DeployID),
                      data = segSample),
                 lmer(meanSpd ~ poly(Tailwind, 2) + Crosswind + (poly(Tailwind, 2)|DeployID),
                      data = segSample),
                 lmer(meanSpd ~ poly(Tailwind, 2) + (1|DeployID),
                      data = segSample),
                 lmer(meanSpd ~ poly(Tailwind, 2) + (poly(Tailwind, 2)|DeployID),
                      data = segSample))
SgTCRank <- data.frame(formula = as.character(lapply(SgTCcand, formula)),
                       AIC = sapply(SgTCcand, AIC)) %>%
  mutate(dAIC = AIC - min(AIC),
         rellik = exp(-0.5 * dAIC),
         AICw = rellik / sum(rellik)) %>%
  arrange(-AICw)
kable(SgTCRank)
SgTCModel <- SgTCcand[[which(SgTCRank$formula[1] == as.character(lapply(SgTCcand, formula)))]]
summary(SgTCModel)
MuMIn::r.squaredGLMM(SgTCModel)

SgAMcand <- list(lmer(meanSpd ~ WindAngle * WindSpd + (1|DeployID),
                      data = segSample),
                 lmer(meanSpd ~ WindAngle * WindSpd + (WindAngle|DeployID),
                      data = segSample),
                 lmer(meanSpd ~ WindAngle + WindSpd + (1|DeployID),
                      data = segSample),
                 lmer(meanSpd ~ WindAngle + WindSpd + (WindAngle|DeployID),
                      data = segSample),
                 lmer(meanSpd ~ WindAngle + (1|DeployID),
                      data = segSample),
                 lmer(meanSpd ~ WindAngle + (WindAngle|DeployID),
                      data = segSample))
SgAMRank <- data.frame(formula = as.character(lapply(SgAMcand, formula)),
                       AIC = sapply(SgAMcand, AIC)) %>%
  mutate(dAIC = AIC - min(AIC),
         rellik = exp(-0.5 * dAIC),
         AICw = rellik / sum(rellik)) %>%
  arrange(-AICw)
kable(SgAMRank)
SgAMModel <- SgAMcand[[which(SgAMRank$formula[1] == as.character(lapply(SgAMcand, formula)))]]
summary(SgAMModel)
MuMIn::r.squaredGLMM(SgAMModel)

SgCand <- c(SgTCcand, SgAMcand)
R2m <- function(m) MuMIn::r.squaredGLMM(m)['R2m']
SgRank <- data.frame(formula = as.character(lapply(SgCand, formula)),
                     r.squared = sapply(SgCand, R2m),
                     AIC = sapply(SgCand, AIC)) %>%
  mutate(dAIC = AIC - min(AIC),
         rellik = exp(-0.5 * dAIC),
         AICw = rellik / sum(rellik)) %>%
  arrange(-AICw) %>%
  select(-rellik)
kable(SgRank)

pred_data <- data.frame(Tailwind = seq(-12, 12, length.out = 100))

SgPred <- function(m) {
  predict(m, newdata = pred_data, re.form = NA)
}
SgBoot <- bootMer(SgTCModel, SgPred, nsim = 1e4, use.u = FALSE, type = 'parametric')
SgCI <- function(bound) {
  if(bound == 'l') {
    quant <- 0.025
  } else {
    quant <- 0.975
  }
  apply(SgBoot$t, 2, function(x) as.numeric(quantile(x, probs = quant, na.rm = TRUE)))
}

p9 <- pred_data %>%
  mutate(PredSpd = predict(SgTCModel, re.form = NA, newdata = .),
         LowerSpd = SgCI('l'),
         UpperSpd = SgCI('u')) %>%
  ggplot(aes(Tailwind, PredSpd)) +
  geom_ribbon(aes(ymin = LowerSpd, ymax = UpperSpd), fill = '#AAAAAA') +
  geom_line() +
  labs(x = 'Tailwind (m/s)',
       y = 'Ground Speed (m/s)') +
  theme_few(base_size = 9)

print(p9)

png('../analysis/figures/Fig9.png', width = 85, height = 85, units = 'mm', res = 300)
print(p9)
dev.off()
```

```{r route_selection}
windspd <- 8
dist <- 75e3

trip_cost <- function(angle) {
  Sout <- predict(SgTCModel, re.form = NA, 
                  newdata = data.frame(Tailwind = windspd * cos(angle)))
  Sin <- predict(SgTCModel, re.form = NA, 
                 newdata = data.frame(Tailwind = windspd * cos(pi - angle)))
  Tout <- dist / Sout
  Tin <- dist / Sin
  Fout <- predict(fam, 
                  newdata = data.frame(WindAngle = angle, 
                                       WindSpd = windspd,
                                       DeployID = 1145),
                  exclude = 's(DeployID)', 
                  type = 'response')
  Fin <- predict(fam, 
                 newdata = data.frame(WindAngle = pi - angle, 
                                      WindSpd = windspd,
                                      DeployID = 1145),
                 exclude = 's(DeployID)', 
                 type = 'response')
  Oout <- predict(oam,
                  newdata = data.frame(WindAngle = angle, 
                                       WindSpd = windspd,
                                       DeployID = 1145),
                  exclude = 's(DeployID)', 
                  type = 'response')
  Oin <- predict(oam,
                 newdata = data.frame(WindAngle = pi - angle, 
                                      WindSpd = windspd,
                                      DeployID = 1145),
                 exclude = 's(DeployID)', 
                 type = 'response')
  EF <- Tout * Fout + Tin * Fin
  EO <- Tout * Oout + Tin * Oin
  data.frame(angle, Sout, Sin, Tout, Tin, Fout, Fin, Oout, Oin, EF, EO)
}

routes0 <- lapply(c(0, pi / 4, pi / 2), trip_cost) %>%
  rbind_all

routes <- routes0 %>%
  group_by(angle) %>%
  summarize(duration = Tout + Tin,
            durHours = duration / 3600, 
            flap = Fout * Tout / duration + Fin * Tin / duration,
            odba = Oout * Tout / duration + Oin * Tin / duration)

kable(routes)

routes %>%
  transmute(angle,
            durHours,
            durRatio = durHours / min(durHours),
            flapRatio = flap / min(flap),
            odbaRatio = odba / min(odba)) %>%
  kable
```
