---
title: "Wind Projection"
author: "Max Czapanskiy"
date: "`r format(Sys.time(), '%d %B, %Y')`"
knit: (function(inputFile, encoding) { 
         out_dir <- '../analysis/reports';
         out_file <- file.path(dirname(inputFile), out_dir, 'WindProjection.html');
         rmarkdown::render(inputFile,
                           encoding = encoding, 
                           output_file = out_file);
       })
output: html_document
---

```{r setup, message = FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	cache = TRUE
)
library(dplyr)
library(lubridate)
library(zoo)
library(ggplot2)
library(foreach)
library(RSQLite)
library(parallel)
library(RNetCDF)
library(oce)
library(iterators)

# dplyr connections
MHI_db <- src_sqlite('../data/MHI_GPS.sqlite')
metadata_db <- tbl(MHI_db, 'Metadata')
tidytracks_db <- tbl(MHI_db, 'TidyTracks')
acceleration_db <- tbl(MHI_db, 'SegmentACC')

POSIX.origin = ymd('1970-01-01', tz = 'UTC')

# Load data from earlier reports
load('../data/out/TransitSegments/Segments.RData')
load('../data/out/TransitSegments/ValidSegments.RData')
load('../data/out/TransitSegments/AccSegments.RData')
```

```{r load_wind}
# File path
wind.path <- '../data/wind/WRF_HI/WRF_Hawaii_Regional_Atmospheric_Model_best.ncd.nc'

# Load NetCDF file
wind.nc <- open.nc(wind.path)

# Print summary
print.nc(wind.nc)

# Dimension names
timeDim <- 'time'   # hours since 2010-05-14 00:00:00.000 UTC
latDim <- 'lat'     # degrees north
lonDim <- 'lon'     # degrees east

# Variable names
uVar <- 'Uwind'
vVar <- 'Vwind'

# Read dimensions and variables
time.wind.origin <- ymd('2010-05-14', tz = 'UTC')
time.wind <- hours(var.get.nc(wind.nc, timeDim)) + time.wind.origin
lat.wind <- var.get.nc(wind.nc, latDim)
lon.wind <- var.get.nc(wind.nc, lonDim)
u.wind <- var.get.nc(wind.nc, uVar)
v.wind <- var.get.nc(wind.nc, vVar)
```

I use approx3d to perform trilinear interpolation to get U and V components approx3d requires equi-spaced dimensions, but the wind model has slight differences in geospatial spacing. These spacing discrepancies are very small, so the wind accessor ends up using re-discretized versions of longitude and latitude dimensions. The time dimension is equi-spaced (1 hour).
```{r wind_spacing}
data.frame(x = lon.wind) %>%
  mutate(xspacing = lead(x) - x) %>%
  ggplot(aes(x = xspacing)) +
  geom_histogram() +
  labs(x = 'Longitude spacing',
       title = 'Difference in x-spacing: 1.53e-5 degrees')
data.frame(y = lat.wind) %>%
  mutate(yspacing = lead(y) - y) %>%
  ggplot(aes(x = yspacing)) +
  geom_histogram() +
  labs(x = 'Latitude spacing',
       title = 'Difference in y-spacing: 1.91e-6 degrees')
```

Functionality for getting wind U and V components
```{r get_wind}
# Re-discretized longitude and latitude
redis.lon <- seq(min(lon.wind), max(lon.wind), length.out = length(lon.wind))
redis.lat <- seq(min(lat.wind), max(lat.wind), length.out = length(lat.wind))
get.wind <- function(component, t, x, y) {
  # Check parameters
  if(length(component) != 1 || !(component %in% c('u', 'v')))
    stop('component parameter must be length 1, "u" or "v"')
  if(!all(between(t, min(time.wind), max(time.wind))))
    stop('t parameter out of bounds')
  
  # Use approx3d from the oce package to perform tri-linear interpolation
  wind.mat <- if(component == 'u') u.wind else v.wind
  # approx3d doesn't work with POSIXct, so t gets re-indexed
  time.wind.int <- seq_along(time.wind)
  t.int <- approx(time.wind, time.wind.int, t)$y
  approx3d(redis.lon, redis.lat, time.wind.int, wind.mat, x, y, t.int)
}

# And a quick test
lon.test <- lon.wind[3]
lat.test <- lat.wind[4]
time.test <- time.wind[5]
u.test <- u.wind[3, 4, 5]
v.test <- v.wind[3, 4, 5]

tol <- 1e-3
if(abs(u.test - get.wind('u', time.test, lon.test, lat.test)) < tol &&
   abs(v.test - get.wind('v', time.test, lon.test, lat.test)) < tol) {
  print('get.wind PASS')
} else {
  print('get.wind FAIL')
}
```

Annotate segments with the wind vectors spatio-temporally closest to the start of the segment. Six segments are too far north and fall outside the wind dataset.
```{r add_uv}
windSegments0 <- accSegments %>%
  mutate(uWind = get.wind('u', Start, Longitude, Latitude),
         vWind = get.wind('v', Start, Longitude, Latitude))

# Six segments fall outside 
windSegments0 %>%
  filter(!between(Longitude, min(lon.wind), max(lon.wind)) | 
           !between(Latitude, min(lat.wind), max(lat.wind)))
```

Project u and v onto flight path to get tail- and crosswind magnitudes
```{r project_wind}
# Function to project wind and get tailwind magnitude
project <- function(b, u, v) {
  b.rad <- b * pi / 180
  x <- cos(b.rad)
  y <- sin(b.rad)
  # U = wind vector, V = flight vector
  # See Methods::Path Analysis
  # U dot V / ||U||
  cbind(u, v, x, y) %>%
    apply(MARGIN = 1, 
          FUN = function(row) {
            U <- row[1:2]
            V <- row[3:4]
            U %*% V
          })
}

# Function to reject wind and get crosswind magnitude (where t is the tailwind magnitude)
reject <- function(u, v, t) {
  cbind(u, v, t) %>%
    apply(MARGIN = 1,
          FUN = function(row) {
            U <- row[1:2]
            t <- row[3]
            sqrt(U %*% U - t^2)
          })
}

# Annotate segments with Tailwind and Crosswind magnitude
windSegments <- windSegments0 %>%
  # Drop the six segments outside the model's spatial range
  na.omit %>%
  mutate(Tailwind = project(bearing, uWind, vWind),
         Crosswind = reject(uWind, vWind, Tailwind))
```

A few plots to visually validate the projection/rejection calculations
```{r proj_valid}
# Verify that tailwind/crosswind magnitude equals wind magnitude
# The histogram of mag([U C]) - mag([T C]) should have a narrow range
windSegments %>%
  # Calculate magnitude of wind two ways: [U V] and [T C]
  mutate(windMag = sqrt(uWind^2 + vWind^2),
         projMag = sqrt(Tailwind^2 + Crosswind^2),
         windProjDiff = windMag - projMag) %>%
  ggplot(aes(x = windProjDiff)) +
  geom_histogram() +
  labs(x = 'Difference between [U V] magnitude and [T C] magnitude (m/s)',
       title = 'Projection/Rejection Validation') +
  scale_x_continuous(labels = function(l) format(l, digits = 2))

# Randomly plot 4 segments with flight vector, wind vector, and rejection/projection
set.seed(1647)
# Vector cross product function from https://stackoverflow.com/questions/15162741/what-is-rs-crossproduct-function/36194961#36194961
# Used for checking if wind vector is left or right of flight vector
vector.cross <- function(a, b) {
    if(length(a)!=3 || length(b)!=3){
        stop("Cross product is only defined for 3D vectors.");
    }
    i1 <- c(2,3,1)
    i2 <- c(3,1,2)
    return (a[i1]*b[i2] - a[i2]*b[i1])
}
foo <- foreach(seg = iter(sample_n(windSegments, 4), by = 'row')) %do% {
  # Rotation matrix, used to make crosswind arrow perpendicular to flight vector
  rot.mat <- matrix(c(0, -1,
                      1, 0),
                    nrow = 2)
  # Flight UNIT vector
  flight.vec <- c(cos(seg$bearing * pi / 180),
                  sin(seg$bearing * pi / 180))
  # Crosswind vector
  crosswind <- seg$Crosswind * rot.mat %*% flight.vec
  # Use the cross product to flip the crosswind as necessary
  if(vector.cross(c(seg$uWind, seg$vWind, 0),
                  c(flight.vec, 0))[3] < 0)
    crosswind <- -crosswind
  
  p <- mutate(seg,
              # Beginning and end of flight vector
              flightX0 = 0,
              flightY0 = 0,
              flightX1 = cos(bearing * pi / 180),
              flightY1 = sin(bearing * pi / 180),
              # Beginning and end of wind vector
              windX0 = 0,
              windY0 = 0,
              windX1 = uWind,
              windY1 = vWind,
              # Beginning and end of tailwind vector
              tailX0 = 0,
              tailY0 = 0,
              tailX1 = Tailwind * flight.vec[1],
              tailY1 = Tailwind * flight.vec[2],
              # Beginning and end of crosswind vector
              crossX0 = tailX1,
              crossY0 = tailY1,
              crossX1 = tailX1 + crosswind[1],
              crossY1 = tailY1 + crosswind[2]) %>%
    ggplot() +
    # Plot tailwind with a dashed line
    geom_segment(aes(x = tailX0,
                     y = tailY0,
                     xend = tailX1,
                     yend = tailY1),
                 arrow = arrow(),
                 linetype = 'dashed') +
    # Plot crosswind with a dashed line
    geom_segment(aes(x = crossX0,
                     y = crossY0,
                     xend = crossX1,
                     yend = crossY1),
                 arrow = arrow(),
                 linetype = 'dashed') +
    # Plot flight vector with a red line
    geom_segment(aes(x = flightX0,
                     y = flightY0,
                     xend = flightX1,
                     yend = flightY1),
                 arrow = arrow(),
                 color = 'red') +
    # Plot wind vector with a blue line
    geom_segment(aes(x = windX0,
                     y = windY0,
                     xend = windX1,
                     yend = windY1),
                 arrow = arrow(),
                 color = 'blue') +
    coord_fixed() +
    labs(x = 'X (m/s)',
         y = 'Y (m/s)',
         title = sprintf('SegmentID %i', seg$SegmentID),
         caption = 'Flight vector in red, wind vector in blue, tail/crosswinds dashed.')
  print(p)
  c(seg$SegmentID)
}
```
